---
title: "HW 12 Population Analysis Pt 2"
author: "Alex Simonyan"
format: pdf
---

## Section 4: Population Scale Analysis

One sample is obviously not enough to know what is happening in a population. You are
interested in assessing genetic differences on a population scale. So, you processed about
~230 samples and did the normalization on a genome level. Now, you want to find whether
there is any association of the 4 asthma-associated SNPs (rs8067378...) on ORMDL3
expression.

>Q13: Read this file into R and determine the sample size for each genotype and their
corresponding median expression levels for each of these genotypes.

How many samples do we have?

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```

```{r}
nrow(expr)
```

```{r}
table(expr$geno)
```

```{r}
library(ggplot2)
```

>Q14:Generate a boxplot with a box per genotype, what could you infer from the relative
expression value between A/A and G/G displayed in this plot? Does the SNP effect the
expression of ORMDL3?

Lets make a boxplot

```{r}
ggplot(expr) + aes(geno, exp, fill=geno) + geom_boxplot(notch=TRUE)
```

G/G expression is the least 
A/A expression is the most 
G/G associated with having a reduced expression of the gene 
Notches don't overlap which means there is very different distributions 
The SNP does effect the expression of ORMDL3









