---
title: "Class 5 Data Visualization with ggplot2"
author: "A16329021"
format: pdf
---

## Using GGPLOT

The ggplot2 package needs to be installed as it does not come with R "out of the box".

We use the `install.packages()` function to do this 
```{r}
head(cars)
```


To use ggplot I need to load it up before I can call any of the functions in the package. I do this with the `library()` function. 

```{r}
library(ggplot2)

```

All ggplot figures have at least 3 things:
- data (the stuff we want to plot)
- aesthetic mapping (aes values)
- geoms

```{r}
ggplot(cars)+ aes(x=speed, y=dist) + geom_point()
```



ggplot is not the only graphing system in R there are lots of others. There is even "base R" graphics. 


```{r}
plot(cars)
```

```{r}
ggplot(cars) + aes(x=speed, y=dist) + geom_point() +geom_smooth(method="lm")
```

```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```

```{r}
nrow(genes)
```

```{r}
colnames(genes)
```
```{r}
ncol(genes)
```
```{r}

```

```{r}
table(genes$State)
```

```{r}
round(table(genes$State) / nrow(genes) *100,2)
```

```{r}
ggplot(genes) + aes(x=Condition1, y=Condition2) + geom_point()
```

```{r}
p <- ggplot(genes) + aes(x=Condition1, y=Condition2, col=State) + geom_point()
p
```

```{r}
p + scale_colour_manual(values = c("blue", "gray", "red"))
```


```{r}
q<- genes
```

```{r}
p <- ggplot(genes) + aes(x=Condition1, y=Condition2, col=State) + geom_point()
p + scale_colour_manual(values = c("blue", "gray", "red")) + labs(title="Gene Expression Changes Upon Drug Treatment", x="Control (no drug)", y="Drug Treatment")
```

```{r}
# File location online
url <- "https://raw.githubusercontent.com/jennybc/gapminder/master/inst/extdata/gapminder.tsv"

gapminder <- read.delim(url)
```
```{r}
# install.packages("dplyr")  ## un-comment to install if needed
library(dplyr)

gapminder_2007 <- gapminder %>% filter(year==2007)
```

```{r}
ggplot(gapminder_2007) + aes(x=gdpPercap, y=lifeExp) + geom_point()
```

```{r}
ggplot(gapminder_2007) +
  aes(x=gdpPercap, y=lifeExp) +
  geom_point(alpha=0.5)
```

```{r}
ggplot(gapminder_2007) +
  aes(x=gdpPercap, y=lifeExp, color=continent, size=pop) +
  geom_point(alpha=0.5)
```

```{r}
ggplot(gapminder_2007) + 
  aes(x = gdpPercap, y = lifeExp, color = pop) +
  geom_point(alpha=0.8)
```

```{r}
ggplot(gapminder_2007) + 
  aes(x = gdpPercap, y = lifeExp, size = pop) +
  geom_point(alpha=0.5)
```

```{r}
ggplot(gapminder_2007) + 
  geom_point(aes(x = gdpPercap, y = lifeExp,
                 size = pop), alpha=0.5) + 
  scale_size_area(max_size = 10)
```

```{r}
gapminder_1957 <- gapminder %>% filter(year==1957)

ggplot(gapminder_1957) + 
  aes(x = gdpPercap, y = lifeExp, color=continent,
                 size = pop) +
  geom_point(alpha=0.7) + 
  scale_size_area(max_size = 10) 
```

```{r}
gapminder_1957 <- gapminder %>% filter(year==1957 | year==2007)

ggplot(gapminder_1957) + 
  geom_point(aes(x = gdpPercap, y = lifeExp, color=continent,
                 size = pop), alpha=0.7) + 
  scale_size_area(max_size = 10) +
  facet_wrap(~year)
```
```{r}
gapminder_top5 <- gapminder %>% 
  filter(year==2007) %>% 
  arrange(desc(pop)) %>% 
  top_n(5, pop)

gapminder_top5
```
```{r}
ggplot(gapminder_top5) + 
  geom_col(aes(x = country, y = pop))
```
```{r}
ggplot(gapminder_top5) + 
  geom_col(aes(x = country, y = pop, fill = continent))
```
```{r}
ggplot(gapminder_top5) + 
  geom_col(aes(x = country, y = pop, fill = lifeExp))
```
```{r}
ggplot(gapminder_top5) +
  aes(x=country, y=pop, fill=gdpPercap) +
  geom_col()
```
```{r}
ggplot(gapminder_top5) +
  aes(x=reorder(country, -pop), y=pop, fill=gdpPercap) +
  geom_col()
```
```{r}
ggplot(gapminder_top5) +
  aes(x=reorder(country, -pop), y=pop, fill=country) +
  geom_col(col="gray30") +
  guides(fill="none")
```
```{r}
head(USArrests)
```

```{r}
USArrests$State <- rownames(USArrests)
ggplot(USArrests) +
  aes(x=reorder(State,Murder), y=Murder) +
  geom_col() +
  coord_flip()
```
```{r}
ggplot(USArrests) +
  aes(x=reorder(State,Murder), y=Murder) +
  geom_point() +
  geom_segment(aes(x=State, 
                   xend=State, 
                   y=0, 
                   yend=Murder), color="blue") +
  coord_flip()
```





